preprocessing

### 2022-11-14 r-prep/prep.Rmd
kaya

data_kaya = data_base %>% 
  group_by(ISO,country,year,GDP_pc,pop,EU,deving_woc,deved,adp_region_woc,adp_region_deving_woc) %>% 
  summarise(CO2=sum(CO2,na.rm=TRUE)) %>%
  left_join(data_iea,by=c("ISO","year")) %>%
  mutate(energy_intensity = energy/GDP_pc) %>%
  mutate(carbon_intensity = CO2/energy) %>%
  na.omit() %>% 
  pivot_longer(c(CO2,pop,GDP_pc,energy_intensity,carbon_intensity),names_to = "var",values_to = "value")

>>>>

data_kaya = data_base %>% 
  group_by(across(-c(sector_title,subsector_title))) %>% 
  summarise(CO2=sum(CO2,na.rm=TRUE)) %>%
  left_join(data_iea,by=c("ISO","year")) %>%
  mutate(energy_intensity = energy/GDP_pc) %>%
  mutate(carbon_intensity = CO2/energy) %>%
  na.omit() %>% 
  pivot_longer(c(CO2,pop,GDP_pc,energy_intensity,carbon_intensity),names_to = "var",values_to = "value")
